Draft final report\\ Measles risk assessment, modelling and benefit--cost analysis\\ \vspace{2 mm} {\large David T S Hayman, Tim Carpenter,\\ Jonathan C Marshall, Mick Roberts, Nigel P French}

### 2014-10-28 draftfinalreport.Rnw
Revisions throughout

\begin{figure}
     \centering
     \includegraphics[width=0.8\textwidth]{incidence_1997_2014.pdf}
     \caption{Measles incidence from 1997 to 2014}
     \label{fig:incidence1997}
\end{figure}

@@

%%%%%%%%%%%%%%%%%%%%%
% INSERT NZ Travlers here
%%%%%%%%%%%%%%%%%%%
% INSERT Total Travelers here
%%%%%%%%%%%%%%%%%%

@@

%%%%%%%%%%%%%%%%%%%
% INSERT NZ risk here
%%%%%%%%%%%%%%%%%%

%%%%%%%%%%%%%%%%%%%
% INSERT Total risk here
%%%%%%%%%%%%%%%%%%
\subsection{Risk analyses methods}
We received the raw EpiSurv measles case data from The Institute of Environmental Science and Research Ltd (ESR) on 27 June 2014. Initial analyses of those ESR data (not shown) suggested that denominator data were required to perform multivariate analyses to avoid confounding results due to a lack of independence among risk factors. Specifically we required $\textsf{Age} \times \textsf{Prioritised Ethnicity} \times \textsf{NZDep}$ data for New Zealand to test whether interactions among case covariates provide additional information on risk over the univariate analyses performed in the \emph{Progress Towards Measles Elimination in New Zealand - Final} report. These $\textsf{Age} \times \textsf{Prioritised Ethnicity} \times \textsf{NZDep}$ data were provided to us on 3 July 2014 by the University of Otago. We used these denominator data to determine if there were interactions among specific age categories, Prioritized Ethnicities, and NZDep that might exist among cases allowing better understanding of risk of measles infection.

@@

@@

\begin{figure}
     \centering
     \includegraphics[width=0.8\textwidth]{naive_map.pdf}
     \caption{Numbers of naive individuals per District Health Board}
     \label{fig:naive_map}
\end{figure}

\begin{figure}
     \centering
     \includegraphics[width=0.8\textwidth]{dhb1.pdf}
     \caption{Numbers of naive individuals per age class, Northland District Health Board}
     \label{fig:Northland}
\end{figure}

\begin{figure}
     \centering
     \includegraphics[width=0.8\textwidth]{dhb2.pdf}
     \caption{Numbers of naive individuals per age class, Waitemata District Health Board}
     \label{fig:Waitemata}
\end{figure}

\begin{figure}
     \centering
     \includegraphics[width=0.8\textwidth]{dhb3.pdf}
     \caption{Numbers of naive individuals per age class, Auckland District Health Board}
     \label{fig:Auckland}
\end{figure}

\begin{figure}
     \centering
     \includegraphics[width=0.8\textwidth]{dhb4.pdf}
     \caption{Numbers of naive individuals per age class, Counties Manukau District Health Board}
     \label{fig:Counties_Manukau}
\end{figure}

\begin{figure}
     \centering
     \includegraphics[width=0.8\textwidth]{dhb5.pdf}
     \caption{Numbers of naive individuals per age class, Waikato District Health Board}
     \label{fig:Waikato}
\end{figure}

\begin{figure}
     \centering
     \includegraphics[width=0.8\textwidth]{dhb6.pdf}
     \caption{Numbers of naive individuals per age class, Lakes District Health Board}
     \label{fig:Lakes}
\end{figure}

\begin{figure}
     \centering
     \includegraphics[width=0.8\textwidth]{dhb7.pdf}
     \caption{Numbers of naive individuals per age class, Bay of Plenty District Health Board}
     \label{fig:BayofPlenty}
\end{figure}

\begin{figure}
     \centering
     \includegraphics[width=0.8\textwidth]{dhb8.pdf}
     \caption{Numbers of naive individuals per age class, Tairawhiti District Health Board}
     \label{fig:Tairawhiti}
\end{figure}

\begin{figure}
     \centering
     \includegraphics[width=0.8\textwidth]{dhb9.pdf}
     \caption{Numbers of naive individuals per age class, Taranaki District Health Board}
     \label{fig:Taranaki}
\end{figure}

\begin{figure}
     \centering
     \includegraphics[width=0.8\textwidth]{dhb10.pdf}
     \caption{Numbers of naive individuals per age class, Hawke's Bay District Health Board}
     \label{fig:HawkesBay}
\end{figure}

\begin{figure}
     \centering
     \includegraphics[width=0.8\textwidth]{dhb11.pdf}
     \caption{Numbers of naive individuals per age class, Whanganui District Health Board}
     \label{fig:Whanganui}
\end{figure}

\begin{figure}
     \centering
     \includegraphics[width=0.8\textwidth]{dhb12.pdf}
     \caption{Numbers of naive individuals per age class, Midcentral District Health Board}
     \label{fig:Midcentral}
\end{figure}

\begin{figure}
     \centering
     \includegraphics[width=0.8\textwidth]{dhb13.pdf}
     \caption{Numbers of naive individuals per age class, Hutt District Health Board}
     \label{fig:Hutt}
\end{figure}

\begin{figure}
     \centering
     \includegraphics[width=0.8\textwidth]{dhb14.pdf}
     \caption{Numbers of naive individuals per age class, Capital and Coast District Health Board}
     \label{fig:CapitalandCoast}
\end{figure}

\begin{figure}
     \centering
     \includegraphics[width=0.8\textwidth]{dhb15.pdf}
     \caption{Numbers of naive individuals per age class, Wairarapa District Health Board}
     \label{fig:Wairarapa}
\end{figure}

\begin{figure}
     \centering
     \includegraphics[width=0.8\textwidth]{dhb16.pdf}
     \caption{Numbers of naive individuals per age class, Nelson Marlborough District Health Board}
     \label{fig:NelsonMarlborough}
\end{figure}

\begin{figure}
     \centering
     \includegraphics[width=0.8\textwidth]{dhb16.pdf}
     \caption{Numbers of naive individuals per age class, West Coast District Health Board}
     \label{fig:WestCoast}
\end{figure}

\begin{figure}
     \centering
     \includegraphics[width=0.8\textwidth]{dhb17.pdf}
     \caption{Numbers of naive individuals per age class, Canterbury District Health Board}
     \label{fig:Canterbury}
\end{figure}

\begin{figure}
     \centering
     \includegraphics[width=0.8\textwidth]{dhb18.pdf}
     \caption{Numbers of naive individuals per age class, South Canterbury District Health Board}
     \label{fig:SouthCanterbury}
\end{figure}

\begin{figure}
     \centering
     \includegraphics[width=0.8\textwidth]{dhb19.pdf}
     \caption{Numbers of naive individuals per age class, Southern District Health Board}
     \label{fig:Southern}
\end{figure}

@@

>>>>

\begin{figure}
     \centering
     \includegraphics[width=1\textwidth]{incidence_1997_2014.pdf}
     \caption{Measles incidence from 1997 to 2014}
     \label{fig:incidence1997}
\end{figure}

@@

%%%%%%%%%%%%%%%%%%
<<echo=FALSE>>=
library(xtable)
totdpdf<-read.csv("mapdata_tot.csv",header=T)
# output top countries
colnames(totdpdf)[1]<-c("country")
toptot <- totdpdf[order(-totdpdf$immigration),] 
immigration<-toptot[1:10,c(1,4)]
topincidence <- totdpdf[order(-totdpdf$incidence),] 
incidence<-topincidence[1:10,c(1,2)]
incidence[,2]<-round(incidence[,2],0)
topvaccine <- totdpdf[order(totdpdf$cover),] 
vaccinecover<-topvaccine[1:10,c(1,3)]
toprisk <- totdpdf[order(-totdpdf$risk),] 
risk<-(toprisk[1:10,c(1,5)])
risk[,2]<-round(risk[,2],0)
@

\begin{table}
\caption{Total travel numbers (2012)}
<<results=tex,echo=FALSE>>=
# This chunk uses "results=tex"
library(Hmisc)
latex(immigration, file="", table.env=FALSE,rowname=NULL)
@
\label{table:immigration12}
\end{table}

%%%%%%%%%%%%%%%%%%%%%

@@

%%%%%%%%%%%%%%%%%%%%%

\begin{table}
\caption{Risk of measles importation to New Zealand in 2012}
<<results=tex,echo=FALSE>>=
# This chunk uses "results=tex"
library(Hmisc)
latex(risk, file="", table.env=FALSE,rowname=NULL)
@
\label{table:risk12}
\end{table}
%%%%%%%%%%%%%%%%%%%
<<echo=FALSE>>=
library(xtable)
nzdpdf<-read.csv("mapdata_nzers.csv",header=T)
# output top countries
colnames(nzdpdf)[1]<-c("country")
topnz <- nzdpdf[order(-nzdpdf$immigration),] 
nz<-topnz[1:10,c(1,4)]
topnzrisk <- nzdpdf[order(-nzdpdf$risk),] 
nzrisk<-(topnzrisk[1:10,c(1,5)])
nzrisk[,2]<-round(nzrisk[,2],0)
@

\begin{table}
\caption{New Zealander travel numbers (2012)}
<<results=tex,echo=FALSE>>=
# This chunk uses "results=tex"
library(Hmisc)
latex(nz, file="", table.env=FALSE,rowname=NULL)
@
\label{table:nztravel12}
\end{table}

%%%%%%%%%%%%%%%%%%%%%

\begin{table}
\caption{Risk of measles importation to New Zealand due to New Zealand travellers in 2012}
<<results=tex,echo=FALSE>>=
# This chunk uses "results=tex"
library(Hmisc)
latex(nzrisk, file="", table.env=FALSE,rowname=NULL)
@
\label{table:nzrisk12}
\end{table}

%%%%%%%%%%%%%%%%%%

%%%%%%%%%%%%%%%%%%
\subsection{Risk analyses methods}
We received the raw EpiSurv measles case data from The Institute of Environmental Science and Research Ltd (ESR) on 27 June 2014. Initial analyses of those ESR data (not shown) suggested that denominator data were required to perform multivariate analyses to avoid confounding results due to a lack of independence among risk factors. Specifically we required $\textsf{Age} \times \textsf{Prioritised Ethnicity} \times \textsf{NZDep}$ data for New Zealand to test whether interactions among case covariates provide additional information on risk over the univariate analyses performed in the \emph{Progress Towards Measles Elimination in New Zealand - Final} report. These $\textsf{Age} \times \textsf{Prioritised Ethnicity} \times \textsf{NZDep}$ data were provided to us on 3 July 2014 by the University of Otago. We used these denominator data to determine if there were interactions among specific age categories, Prioritized Ethnicities, and NZDep that might exist among cases allowing better understanding of risk of measles infection.

@@

\begin{figure}
     \begin{center}
     \includegraphics[width=1\textwidth]{nzers.pdf}
     \end{center}
     \caption{Trends in travel}
     \label{fig:travel}
\end{figure}

@@

\begin{figure}
     \begin{center}
     \includegraphics[width=0.6\textwidth]{naive_map.pdf}
     \end{center}
     \caption{Numbers of naive individuals per District Health Board}
     \label{fig:naive_map}
\end{figure}

\begin{figure}
     \begin{center}
     \includegraphics[width=0.6\textwidth]{dhb1.pdf}
     \end{center}
     \caption{Numbers of naive individuals per age class, Northland District Health Board}
     \label{fig:Northland}
\end{figure}

\begin{figure}
     \begin{center}
     \includegraphics[width=0.6\textwidth]{dhb2.pdf}
     \end{center}
     \caption{Numbers of naive individuals per age class, Waitemata District Health Board}
     \label{fig:Waitemata}
\end{figure}

\begin{figure}
     \begin{center}
     \includegraphics[width=0.6\textwidth]{dhb3.pdf}
     \end{center}
     \caption{Numbers of naive individuals per age class, Auckland District Health Board}
     \label{fig:Auckland}
\end{figure}

\begin{figure}
     \begin{center}
     \includegraphics[width=0.6\textwidth]{dhb4.pdf}
     \end{center}
     \caption{Numbers of naive individuals per age class, Counties Manukau District Health Board}
     \label{fig:Counties_Manukau}
\end{figure}

\begin{figure}
     \begin{center}
     \includegraphics[width=0.6\textwidth]{dhb5.pdf}
     \end{center}
     \caption{Numbers of naive individuals per age class, Waikato District Health Board}
     \label{fig:Waikato}
\end{figure}

\begin{figure}
     \begin{center}
     \includegraphics[width=0.6\textwidth]{dhb6.pdf}
     \end{center}
     \caption{Numbers of naive individuals per age class, Lakes District Health Board}
     \label{fig:Lakes}
\end{figure}

\begin{figure}
     \begin{center}
     \includegraphics[width=0.6\textwidth]{dhb7.pdf}
     \end{center}
     \caption{Numbers of naive individuals per age class, Bay of Plenty District Health Board}
     \label{fig:BayofPlenty}
\end{figure}

\begin{figure}
     \begin{center}
     \includegraphics[width=0.6\textwidth]{dhb8.pdf}
     \end{center}
     \caption{Numbers of naive individuals per age class, Tairawhiti District Health Board}
     \label{fig:Tairawhiti}
\end{figure}

\begin{figure}
     \begin{center}
     \includegraphics[width=0.6\textwidth]{dhb9.pdf}
     \end{center}
     \caption{Numbers of naive individuals per age class, Taranaki District Health Board}
     \label{fig:Taranaki}
\end{figure}

\begin{figure}
     \begin{center}
     \includegraphics[width=0.6\textwidth]{dhb10.pdf}
     \end{center}
     \caption{Numbers of naive individuals per age class, Hawke's Bay District Health Board}
     \label{fig:HawkesBay}
\end{figure}

\begin{figure}
     \begin{center}
     \includegraphics[width=0.6\textwidth]{dhb11.pdf}
     \end{center}
     \caption{Numbers of naive individuals per age class, Whanganui District Health Board}
     \label{fig:Whanganui}
\end{figure}

\begin{figure}
     \begin{center}
     \includegraphics[width=0.6\textwidth]{dhb12.pdf}
     \end{center}
     \caption{Numbers of naive individuals per age class, Midcentral District Health Board}
     \label{fig:Midcentral}
\end{figure}

\begin{figure}
     \begin{center}
     \includegraphics[width=0.6\textwidth]{dhb13.pdf}
     \end{center}
     \caption{Numbers of naive individuals per age class, Hutt District Health Board}
     \label{fig:Hutt}
\end{figure}

\begin{figure}
     \begin{center}
     \includegraphics[width=0.6\textwidth]{dhb14.pdf}
     \end{center}
     \caption{Numbers of naive individuals per age class, Capital and Coast District Health Board}
     \label{fig:CapitalandCoast}
\end{figure}

\begin{figure}
     \begin{center}
     \includegraphics[width=0.6\textwidth]{dhb15.pdf}
     \end{center}
     \caption{Numbers of naive individuals per age class, Wairarapa District Health Board}
     \label{fig:Wairarapa}
\end{figure}

\begin{figure}
     \begin{center}
     \includegraphics[width=0.6\textwidth]{dhb16.pdf}
     \end{center}
     \caption{Numbers of naive individuals per age class, Nelson Marlborough District Health Board}
     \label{fig:NelsonMarlborough}
\end{figure}

\begin{figure}
     \begin{center}
     \includegraphics[width=0.6\textwidth]{dhb17.pdf}
     \end{center}
     \caption{Numbers of naive individuals per age class, West Coast District Health Board}
     \label{fig:WestCoast}
\end{figure}

\begin{figure}
     \begin{center}
     \includegraphics[width=0.6\textwidth]{dhb18.pdf}
     \end{center}
     \caption{Numbers of naive individuals per age class, Canterbury District Health Board}
     \label{fig:Canterbury}
\end{figure}

\begin{figure}
     \begin{center}
     \includegraphics[width=0.6\textwidth]{dhb19.pdf}
     \end{center}
     \caption{Numbers of naive individuals per age class, South Canterbury District Health Board}
     \label{fig:SouthCanterbury}
\end{figure}

\begin{figure}
     \begin{center}
     \includegraphics[width=0.6\textwidth]{dhb20.pdf}
     \end{center}
     \caption{Numbers of naive individuals per age class, Southern District Health Board}
     \label{fig:Southern}
\end{figure}

@@

<<echo=FALSE>>=
costtable20<-read.csv("cost_benefit_20.csv",header=T)
@

\begin{sidewaystable}\small
\caption{Cost benefit analyses with 20 dollars per vaccine}
<<results=tex,echo=FALSE>>=
# This chunk uses "results=tex"
library(Hmisc)
latex(costtable20, file="", table.env=FALSE,rowname=NULL)
@
\label{table:cost20}
\end{sidewaystable}

<<echo=FALSE>>=
costtable50<-read.csv("cost_benefit_50.csv",header=T)
@

\begin{sidewaystable}\small
\caption{Cost benefit analyses with 50 dollars per vaccine}
<<results=tex,echo=FALSE>>=
# This chunk uses "results=tex"
library(Hmisc)
latex(costtable50, file="", table.env=FALSE,rowname=NULL)
@
\label{table:cost50}
\end{sidewaystable}
